Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-14 writing/embedded_meta_learning.tex
Beginning to incorporate Erin's comments

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, meta-mappings, which involve systematically transforming task behaviors. We suggest that the key to achieving this flexibility is representing the task being performed along with the computations used to perform it. We therfore draw inspiration from meta-learning and functional programming to propose a class of architectures that represent both data and tasks in a shared latent space. We demonstrate the flexibility of these architectures by showing that they can perform meta-mappings, i.e. that they can exhibit zero-shot remapping of behavior on new tasks. 
\end{abstract}

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing demands. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits, such as a useful framework for continual learning. We suggest that approaches like this will be key to building more intelligent and flexible deep learning systems. \par

\section{Meta-mappings}
We propose the challenge of meta-mappings, that is, mappings where an input or output (or both) is a task. These can take several flavors, including mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that these meta-mappings are a useful way to think about the flexibility that humans possess. For example, a system that learns to play go should be able to adapt to a different board size. These tasks clearly share a great deal of structure, and a human can easily transfer much of their knowledge from one board size to another without retraining, but deep learning models generally have no way of achieving this. Similarly, suppose an agent knows how to navigate from the airport to downtown. It should be able to adapt to changes of the environment such as street closures. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not capture the extent of human flexibility. In particular, we can use our knowledge of a task to adapt to these type of task alterations zero-shot, that is, without seeing a single example from the new task. \par
In the machine-learning setting, these mappings can be cued in different ways, for example by examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to change your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board''). Achieving this kind of flexibility will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

>>>>

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? Toward this goal, we propose a new class of challenges, and a class of architectures that can solve them. The challenges are in the form of meta-mappings, which involve systematically transforming task behaviors to adapt to new tasks. We suggest that the key to achieving these challenges is representing the task being performed along with the computations used to perform it. We therefore draw inspiration from meta-learning and functional programming to propose a class of Embedded Meta-Learning (EML) architectures that represent both data and tasks in a shared latent space. We demonstrate the flexibility of these architectures by showing that they can perform meta-mappings, i.e. that they can exhibit zero-shot remapping of behavior on new tasks. 
\end{abstract}

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can use our knowledge of a task flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing demands. We demonstrate that our arhitectures can flexibly remap their behavior to address the meta-mapping challenge. EML may also offer other benefits, such as a useful framework for continual learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

\section{Meta-mappings}
We propose the challenge of meta-mappings, that is, tasks which take tasks as an input or output. These can take several flavors, including mapping from tasks to language (explaining), mapping from language to tasks (following instructions), and mapping from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that these meta-mappings are a useful way to think about the flexibility that humans possess. For example, a system that learns to play go should be able to adapt to a different board size. Go on a large board or a small board share a great deal of structure, and a human can easily transfer much of their knowledge from one size to another without retraining, but deep learning models generally have no way of achieving this. Similarly, suppose an agent knows how to navigate from the airport to downtown. It should be able to adapt to changes of the environment such as street closures. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not capture the extent of human flexibility. In particular, we can use our knowledge of a task to adapt to task alterations zero-shot, that is, without seeing a single example from the new task. \par
Meta-mappings can be cued in different ways, for example by examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to change your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board''). Achieving this kind of flexibility will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par
